---
title: "Notes & planning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We can use this document to take notes & plan.




2/1/2021


Phenology timeseries modeling ideas
* First goal: identify a couple important variables and implement a simple model


Variables
* What are we getting with the incoming data?
* How do we want to represent time/seasonality?
  * Parameterize it in terms of change (i.e. delta-gcc)?
  * Smooth function (e.g. spline / GAM)?
  * Additive effect of mean of historical data for that date?
*





One idea: additive components. From Hyndman and Athanasopoulos book:
$$y_t = \mathcal{N} (S_t + T_t + R_t, \sigma_y)$$
$$R_t = x_t^T\beta$$

where S_t is the seasonal component, T_t is the trend component, and R_t is the
"remainder," or what I'd call the anomaly.

We might say that T_t is 0 when we've already accounted for the seasonal and temporal trend (maybe, maybe not).

Maybe S_t is driven by some values we're interested in (e.g. Julian date) but we
could also say that $S_t = S_{jd}$ in every year.


```{r}
nimbleCode({
  
  for (t in 1:ntime) {
    for (site in 1:nsite) {
      
      y[site, t] ~ dnorm(
        mean = St[site, date_of_year[t]],
        sd = sd_y
      )
      
    }
  }
})
```


# Data
NEON meteorological station and NOAA’s Global Ensemble Forecast System (GEFS) will be provided. NOAA Global Ensemble Forecasting System 35-day weather forecasts for each NEON site: 3-6 hourly (https://data.ecoforecast.org/minio/drivers/noaa/NOAAGEFS_6hr/) and hourly (https://data.ecoforecast.org/minio/drivers/noaa/NOAAGEFS_1hr/) 



